05_DESeq_manipulation

### 2019-05-13 analysis/05_DESeq2_manipulate.Rmd
now with fewer groupings

# set levels
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("m.inc.d3",  "m.inc.d9",
                                "m.inc.d17", "m.inc.d8", "prolong", "extend", "m.n2"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
```

@@

```{r althings}
for (eachgroup in levels(m.colData$sextissue)){

@@

dat$V1 <- factor(dat$V1, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))
dat$V2 <- factor(dat$V2, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))

@@

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- NULL
colnames(sampleDistMatrix) <- colData$treatment
colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         fontsize = 6, 
         main = eachgroup)
}
```

>>>>

# set levels
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("m.inc.d3",  "m.inc.d9",
                                "m.inc.d17", "m.n2",
                                "m.inc.d8", "prolong", "extend"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))

m.colData$outcome <- ifelse(grepl("d3|d9|d17|n2", m.colData$treatment), "remove", 
                     ifelse(grepl("d8", m.colData$treatment),"early",
                     ifelse(grepl("prolong", m.colData$treatment),"prolong",
                     ifelse(grepl("extend", m.colData$treatment),"extend", NA))))
m.colData$outcome <- factor(m.colData$outcome, levels = 
                              c("remove",  "prolong",
                                "early", "extend"))
summary(m.colData[c(7,3,4,5,9)])
```

@@

```{r althings, eval = FALSE}

for (eachgroup in levels(m.colData$sextissue)){

@@

dat$V1 <- factor(dat$V1, levels = 
                              c("m.inc.d3",  "m.inc.d9",
                                "m.inc.d17", "m.n2",
                                "m.inc.d8", "prolong", "extend"))
dat$V2 <- factor(dat$V2, levels = 
                              c("m.inc.d3",  "m.inc.d9",
                                "m.inc.d17", "m.n2",
                                "m.inc.d8", "prolong", "extend"))

@@

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- NULL
colnames(sampleDistMatrix) <- colData$treatment
colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)

#pheatmap(sampleDistMatrix,
#         clustering_distance_rows=sampleDists,
#         clustering_distance_cols=sampleDists,
#         col=colors,
 #        fontsize = 6, 
 #        main = eachgroup)
}
```

```{r allthings-outcome}

for (eachgroup in levels(m.colData$sextissue)){

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ outcome )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized 
head(vsd)

numDEGs <- function(group1, group2){
  res <- results(dds, contrast = c("outcome", group1, group2), independentFiltering = T)
  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
  return(sumpadj)}

#create list of groups
a <- levels(colData$outcome)
b <- levels(colData$outcome)

dat$V1 <- factor(dat$V1, levels = 
                              c("remove",  "prolong",
                                "early", "extend"))
dat$V2 <- factor(dat$V2, levels = 
                              c("remove",  "prolong",
                                "early", "extend"))

allcontrasts <- dat %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#FFFFFF00") + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("outcome"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

print(summary(aov(PC1 ~ outcome, data=pcadata)))
print(TukeyHSD(aov(PC1 ~ outcome, data=pcadata), which = "outcome"))

print(summary(aov(PC2 ~ outcome, data=pcadata))) 
print(TukeyHSD(aov(PC2 ~ outcome, data=pcadata), which = "outcome"))

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
#print(pca34)
}
```
